### 2018-07-23 contents/sagit.tex
Updates from ShareLaTeX
final sync, 2-sided print correction

\subsection{Preprocessing}
In order to automate the generation of tractography using multiple methods and minimize user-bias, the SAGIT framework is created (Figure 1). DWI sequences were corrected for eddy-current and motion distortions with appropriate rotational corrections to gradient vectors \cite{Leemans2009}. Fractional anisotropy (FA), axial diffusivity (AD) and radial diffusivity (RD) maps were generated from the DWI. A group specific average anatomical template was created with subject T$_1$ image, and intra-subject T$_1$ to DWI space was obtained using symmetric diffeomorphic registration (SyN) with Automated Normalization Tools (ANTs) \cite{Avants2008b}. Cortical and subcortical segmentations based on the T$_1$ images were obtained with FreeSurfer segmentation software \cite{Fischl2002b}.

@@

\subsection{Tractography Evaluation}
The generated visual reports were then evaluated by 5 neuroscientists in a blinded study using a set of rating criteria (See supplemental). The conjunction images were at the same time used to generate NOS scores. The NOS score quantified the conjunction image generated by overlapping tractography masks to meaningfully determine the spatial agreement independent of anatomy. The assumption was that given a conjunction image, where its voxel value s denotes a range of overlap percentages between 0\% to 100\%, and value 1 that denotes the 100\% overlap value, we formalize it as $ s\in[0,1]$, and define NOS score as: 
\begin{equation}
n^{-1} \sum_{i=0}^{n-1} \frac{\ln(v_i)}{\ln(v_0)}
\end{equation}
where $n$ was the number of bins; $v_0$ was the number of voxels where $s > 0$;  $v_i$ was the number of voxels with $ 0 < s \ge \frac{i}{n}$; this study used $n=10$. Examples of NOS score behaviour are available in Figure 4. 
The rater scores were then normalized to each anatomy, and the normalized rater scores were linearly regressed against the NOS scores for correlations. Linear regression of rater scores against NOS was performed using the R-statistics library [R Development Core Team, 2012].

\begin{figure}[p]
\centering
\includegraphics[width=140mm]{figure1.pdf}
\caption{Flow diagram of the SAGIT group tractography framework. The SAGIT framework is designed to be fully configurable and extensible.}
\label{fig:SGfigure1}
\end{figure}

@@

The fornix was shown to be consistent across the different algorithms. The visual report (Figure 2; the complete visual report can be found in supplemental) suggested that the subregions of the fornix were consistently delineated. The merged fornix tracts (Figure 3) showed different patterns of streamline distributions. CST$_{prob}$ and XST both showed wide spread streamline dispersions, while CST$_{det}$ and particularly DTT were limited to the region of the fornix. The fornix ratings (Figure 5) showed essentially no variability between raters.  The fornix NOS scores (Figure 6) similarly were consistent across the algorithms with little variability. Both filtered and unfiltered ratings were identical.

For the CN VII/VIII, CST$_{prob}$, CST$_{det}$ and XST were all able to delineate the cranial as well as brainstem segments of the nerve (Figure 2). CST$_{prob}$ however also delineated much of the cerebellum. DTT by Slicer was not able to delineate the brainstem portions of the fibre and therefore was scored lower than the other methods. These differences in delineation can be more clearly seen in the merged tracts (Figure 3), where the XST also showed further lateral delineations of CN VII/VIII. In unfiltered tractography, there was much more inter-rater variability for all except CST$_{prob}$. CST$_{det}$ and XST also showed lower average rating when unfiltered (Figure 5). The NOS showed similar trend where DTT resulted in the lowest score (Figure 6). The NOS also showed higher score for CST$_{det}$ when comparing to XST. There were no notable changes in NOS scores in unfiltered tractography.

For CN X, CST$_{prob}$ was rated the highest with no inter-rater variability. CST$_{det}$ and XST resulted in more rating variability and a lower average score (Figure 5). CST$_{prob}$ visually showed delineations that extended into the ipsilateral higher brain regions even with filters (Figures 2, 3). DTT showed the worst reproducibility, although its anatomy was recognizable in a small number of individuals.  In unfiltered tractography, CST$_{det}$ and XST showed reduced variability, and visually showed delineations of ascending and brainstem projects. DTT also showed improved ratings when unfiltered. NOS scores (Figure 6) for both filtered and unfiltered CN X tractography similarly showed an overall increase in score with unfiltered tractography.

For the red nucleus projections, visually CST$_{prob}$ showed the most consistent reproducibility of the decussation (Figure 2, Figure 3), with XST showing higher reproducibility over CST$_{det}$. DTT failed to delineate the pathway and instead delineated an erroneous path that decussated at the corpus callosum. CST$_{prob}$, CST$_{det}$ and XST resulted similar ratings, with unfiltered tracts showing lower rating and higher variability. DTT had the lowest rating, and very high variability (Figure 5). NOS score of the red nucleus projections highlighted the stronger performance of the CST$_{prob}$ over the other algorithms, and XST result was scored higher than CST$_{det}$ (Figure 6). The NOS scores of unfiltered tracts were notably higher for all algorithms, as the resulting tracts seemed to delineate much of the cortical spinal projections.

 The optic radiation showed the greatest reproducibility with CST$_{prob}$, where the Meyer’s loop could be seen in the result (Figure 2, Figure 3). CST$_{det}$ and XST showed low reproducibility of less than 20\%, and DTT failed to delineate any structure on the right side (see supplemental). The ratings reflected the observation (Figure 5), with CST$_{prob}$ having the full rating. CST$_{det}$ was rated higher in filtered tractography, while XST was rated higher when unfiltered. DTT also showed higher ratings when unfiltered. NOS score characterized the drop-off in reproducibility in filtered tractography for the deterministic methods (Figure 6), and the trend of rating under filtered tractography matched well with the ratings. NOS scores for unfiltered tractography however showed little match with ratings.

The auditory radiation was the least reproducible region. In filtered tractography, although CST$_{prob}$ showed high reproducibility in tractography (Figure 2, Figure 3), it was not clear if its delineation was correct. Visually there were wide spread area of false positives towards the occipital and anterior temporal lobe. The deterministic methods all showed poor ability to reach the Heschl’s gyrus. DTT was completely unable to produce any structures at all. For filtered ratings,  CST$_{prob}$ was the highest, and with XST rated higher than CST$_{det}$ (Figure 5). DTT scored zero due to the lack of delineated structures. When unfiltered, all algorithms showed widespread false positives in the hemispheres. Unfiltered ratings were over all higher than filtered ratings. NOS scores’ trends for filtered tractography matched well with ratings, but also highlighted the large differences in reproducibility (Figure 6). While NOS scores under unfiltered tractography did not correlate with ratings.

Linear regression of rater scores against NOS (Figure 7) showed significant (p\textless0.05) correlation of the two sets of scores in filtered tractography. However correlations were not significant (p\textgreater0.05) for unfiltered tractography. It can be observed that NOS score in unfiltered tractography agrees well with visual intuition, however rater variability and anatomical reports suggested greater false positives when presented with unfiltered results.

@@

\afterpage{%
    \clearpage% Flush earlier floats (otherwise order might not be correct)
    \thispagestyle{empty}% empty page style (?)
    \begin{landscape}% Landscape page
\begin{figure}[p]
\includegraphics[width=\linewidth]{figure3.png}
\caption[Example of merged tractography models for each of the techniques and anatomy.]{Example of merged tractography models for each of the techniques and anatomy. The fornix is coloured to distinguish tracts from each of the 42 subjects, in order to demonstrate the effect of combining the group data. The other anatomies are coloured by FA intensity. The specific FA measures from each point were sampled from their native DWI space before tractography deformation. The grouping of the similar FA measures across the subjects show that the registrations were accurate. The Red Nuclei shows different metrics (FA and AD) depending on lateralization, in order to highlight the pontine decussation.}
\centering
\label{fig:SGfigure3}
\end{figure}
    \end{landscape}
    \clearpage% Flush page
}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\textwidth]{figure5.pdf}
	\caption[Normalized anatomical scores as rated by experts.]{Normalized anatomical scores as rated by experts. The error bars denote standard deviations of the ratings. The fornix serves as the control anatomy, therefore the represented filtered and unfiltered image are identical. It can be observed that unfiltered anatomy increases rater variability in some and decrease in others.}
	\label{fig:SGfigure5}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figure6.pdf}
\caption{NOS scores for each anatomy. It can be noted that having ROI filters severely reduces reproducibility in deterministic methods.}
\label{fig:SGfigure6}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figure7.pdf}
\caption{Correlation between rating scores and NOS. Filtered tractography NOS is significantly (p\textless 0.05) correlated with ratings. However unfiltered tractography show no significance.}
\label{fig:SGfigure7}
\end{figure}

@@

When attempting to judge tractography delineation quality, human experts often are able to assess a particular delineation by visually identifying known anatomical priors. Direct tractography visualizations suggest that wide streamline dispersions often obscure anatomical details in visualizations, making anatomical assessment difficult (figure 3). Therefore, image conjunctions offer better ability to assess reproducibility than direct tractography renderings alone. We aimed to develop an automated assessment of tractography reproducibility, since a) researchers are often prone to judgment bias when exposed to large number of tractography results, and that subtle differences in results are often hard to distinguish, and b) in an automated pipeline, the assessment score is necessary to allow iteratively fine tuning of tractography parameters. Since current neuroimaging techniques are not yet able to reliably classify neuroanatomy from tractography or image-based morphology, it is desirable for a computing pipeline to optimize and present the most reproducible tractography delineating across a population based on a particular set of ROIs as a first step. Subsequent anatomical assessment can then be made by a human expert, and the appropriate changes in ROI filter strategy can be made. This approach allows researchers with strong neuroanatomy backgrounds, but are less technically inclined, to make better judgments when performing group tractography.

@@

The limit of the NOS score is closed tied to the performance of its associated tractography algorithm. An algorithm and its associated parameters may consistently produce the wrong result in all subjects and produce a high NOS score, or it may produce no results, and result in a NOS score of 0. Both of these cases are regular occurrences in single-subject tractography analysis. With merged tractography, it is easier to recognize such faults. This is because in practice, incorrect tract path are often the result of imaging anomalies or algorithm limit, and are unstable across a population. This can be observed as a general trend in Figure 3, for example where DTI delineation of the rubrocerebellar tract is clearly incorrect, and results in low NOS score. Therefore, NOS score is capable of characterizing the uncertainty as a result of low algorithmic performance. In the case of false positives that are highly consistent, SAGIT can help researchers making more informed judgement.

>>>>

\subsection{Preprocessing}
In order to automate the generation of tractography using multiple methods and minimize user-bias, the SAGIT framework is created (Figure \ref{fig:SGfigure1}). DWI sequences were corrected for eddy-current and motion distortions with appropriate rotational corrections to gradient vectors \cite{Leemans2009}. Fractional anisotropy (FA), axial diffusivity (AD) and radial diffusivity (RD) maps were generated from the DWI. A group specific average anatomical template was created with subject T$_1$ image, and intra-subject T$_1$ to DWI space was obtained using symmetric diffeomorphic registration (SyN) with Automated Normalization Tools (ANTs) \cite{Avants2008b}. Cortical and subcortical segmentations based on the T$_1$ images were obtained with FreeSurfer segmentation software \cite{Fischl2002b}.

@@

\subsection{Tractography Evaluation}
The generated visual reports were then evaluated by 5 neuroscientists in a blinded study using a set of rating criteria (See supplemental). The conjunction images were at the same time used to generate NOS scores. The NOS score quantified the conjunction image generated by overlapping tractography masks to meaningfully determine the spatial agreement independent of anatomy. The assumption was that given a conjunction image, where its voxel value s denotes a range of overlap percentages between 0\% to 100\%, and value 1 that denotes the 100\% overlap value, we formalize it as $ s\in[0,1]$, and define NOS score as: 
\begin{equation}
n^{-1} \sum_{i=0}^{n-1} \frac{\ln(v_i)}{\ln(v_0)}
\end{equation}
where $n$ was the number of bins; $v_0$ was the number of voxels where $s > 0$;  $v_i$ was the number of voxels with $ 0 < s \ge \frac{i}{n}$; this study used $n=10$. Examples of NOS score behaviour are available in Figure \ref{fig:SGfigure4}. 
The rater scores were then normalized to each anatomy, and the normalized rater scores were linearly regressed against the NOS scores for correlations. Linear regression of rater scores against NOS was performed using the R-statistics library [R Development Core Team, 2012].

\begin{figure}[p]
\centering
\includegraphics[width=0.8\linewidth]{figure1.pdf}
\caption{Flow diagram of the SAGIT group tractography framework. The SAGIT framework is designed to be fully configurable and extensible.}
\label{fig:SGfigure1}
\end{figure}

@@

The fornix was shown to be consistent across the different algorithms. The visual report (Figure \ref{fig:SGfigure2}; the complete visual report can be found in supplemental) suggested that the subregions of the fornix were consistently delineated. The merged fornix tracts (Figure \ref{fig:SGfigure3}) showed different patterns of streamline distributions. CST$_{prob}$ and XST both showed wide spread streamline dispersions, while CST$_{det}$ and particularly DTT were limited to the region of the fornix. The fornix ratings (Figure \ref{fig:SGfigure5}) showed essentially no variability between raters.  The fornix NOS scores (Figure \ref{fig:SGfigure6}) similarly were consistent across the algorithms with little variability. Both filtered and unfiltered ratings were identical.

For the CN VII/VIII, CST$_{prob}$, CST$_{det}$ and XST were all able to delineate the cranial as well as brainstem segments of the nerve (Figure \ref{fig:SGfigure2}). CST$_{prob}$ however also delineated much of the cerebellum. DTT by Slicer was not able to delineate the brainstem portions of the fibre and therefore was scored lower than the other methods. These differences in delineation can be more clearly seen in the merged tracts (Figure \ref{fig:SGfigure3}), where the XST also showed further lateral delineations of CN VII/VIII. In unfiltered tractography, there was much more inter-rater variability for all except CST$_{prob}$. CST$_{det}$ and XST also showed lower average rating when unfiltered (Figure \ref{fig:SGfigure5}). The NOS showed similar trend where DTT resulted in the lowest score (Figure \ref{fig:SGfigure6}). The NOS also showed higher score for CST$_{det}$ when comparing to XST. There were no notable changes in NOS scores in unfiltered tractography.

For CN X, CST$_{prob}$ was rated the highest with no inter-rater variability. CST$_{det}$ and XST resulted in more rating variability and a lower average score (Figure \ref{fig:SGfigure5}). CST$_{prob}$ visually showed delineations that extended into the ipsilateral higher brain regions even with filters (Figures \ref{fig:SGfigure2}, \ref{fig:SGfigure3}). DTT showed the worst reproducibility, although its anatomy was recognizable in a small number of individuals.  In unfiltered tractography, CST$_{det}$ and XST showed reduced variability, and visually showed delineations of ascending and brainstem projects. DTT also showed improved ratings when unfiltered. NOS scores (Figure \ref{fig:SGfigure6}) for both filtered and unfiltered CN X tractography similarly showed an overall increase in score with unfiltered tractography.

For the red nucleus projections, visually CST$_{prob}$ showed the most consistent reproducibility of the decussation (Figure \ref{fig:SGfigure2}, Figure \ref{fig:SGfigure3}), with XST showing higher reproducibility over CST$_{det}$. DTT failed to delineate the pathway and instead delineated an erroneous path that decussated at the corpus callosum. CST$_{prob}$, CST$_{det}$ and XST resulted similar ratings, with unfiltered tracts showing lower rating and higher variability. DTT had the lowest rating, and very high variability (Figure \ref{fig:SGfigure5}). NOS score of the red nucleus projections highlighted the stronger performance of the CST$_{prob}$ over the other algorithms, and XST result was scored higher than CST$_{det}$ (Figure \ref{fig:SGfigure6}). The NOS scores of unfiltered tracts were notably higher for all algorithms, as the resulting tracts seemed to delineate much of the cortical spinal projections.

 The optic radiation showed the greatest reproducibility with CST$_{prob}$, where the Meyer’s loop could be seen in the result (Figure \ref{fig:SGfigure2}, Figure \ref{fig:SGfigure3}). CST$_{det}$ and XST showed low reproducibility of less than 20\%, and DTT failed to delineate any structure on the right side (see supplemental). The ratings reflected the observation (Figure \ref{fig:SGfigure5}), with CST$_{prob}$ having the full rating. CST$_{det}$ was rated higher in filtered tractography, while XST was rated higher when unfiltered. DTT also showed higher ratings when unfiltered. NOS score characterized the drop-off in reproducibility in filtered tractography for the deterministic methods (Figure \ref{fig:SGfigure6}), and the trend of rating under filtered tractography matched well with the ratings. NOS scores for unfiltered tractography however showed little match with ratings.

The auditory radiation was the least reproducible region. In filtered tractography, although CST$_{prob}$ showed high reproducibility in tractography (Figure \ref{fig:SGfigure2}, Figure \ref{fig:SGfigure3}), it was not clear if its delineation was correct. Visually there were wide spread area of false positives towards the occipital and anterior temporal lobe. The deterministic methods all showed poor ability to reach the Heschl’s gyrus. DTT was completely unable to produce any structures at all. For filtered ratings,  CST$_{prob}$ was the highest, and with XST rated higher than CST$_{det}$ (Figure \ref{fig:SGfigure5}). DTT scored zero due to the lack of delineated structures. When unfiltered, all algorithms showed widespread false positives in the hemispheres. Unfiltered ratings were over all higher than filtered ratings. NOS scores’ trends for filtered tractography matched well with ratings, but also highlighted the large differences in reproducibility (Figure \ref{fig:SGfigure6}). While NOS scores under unfiltered tractography did not correlate with ratings.

Linear regression of rater scores against NOS (Figure \ref{fig:SGfigure7}) showed significant (p\textless0.05) correlation of the two sets of scores in filtered tractography. However correlations were not significant (p\textgreater0.05) for unfiltered tractography. It can be observed that NOS score in unfiltered tractography agrees well with visual intuition, however rater variability and anatomical reports suggested greater false positives when presented with unfiltered results.

@@

\afterpage{%
    \clearpage% Flush earlier floats (otherwise order might not be correct)
    \thispagestyle{empty}% empty page style (?)
    \begin{landscape}% Landscape page
\begin{figure}[p]
\centering
\includegraphics[width=0.9\linewidth]{figure3.png}
\caption[Example of merged tractography models for each of the techniques and anatomy.]{Example of merged tractography models for each of the techniques and anatomy. The fornix is coloured to distinguish tracts from each of the 42 subjects, in order to demonstrate the effect of combining the group data. The other anatomies are coloured by FA intensity. The specific FA measures from each point were sampled from their native DWI space before tractography deformation. The grouping of the similar FA measures across the subjects show that the registrations were accurate. The Red Nuclei shows different metrics (FA and AD) depending on lateralization, in order to highlight the pontine decussation.}
\label{fig:SGfigure3}
\end{figure}
    \end{landscape}
    \clearpage% Flush page
}

\begin{figure}[ht]
	\centering
	\includegraphics[width=\linewidth]{figure5.pdf}
	\caption[Normalized anatomical scores as rated by experts.]{Normalized anatomical scores as rated by experts. The error bars denote standard deviations of the ratings. The fornix serves as the control anatomy, therefore the represented filtered and unfiltered image are identical. It can be observed that unfiltered anatomy increases rater variability in some and decrease in others.}
	\label{fig:SGfigure5}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure6.pdf}
\caption{NOS scores for each anatomy. It can be noted that having ROI filters severely reduces reproducibility in deterministic methods.}
\label{fig:SGfigure6}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{figure7.pdf}
\caption{Correlation between rating scores and NOS. Filtered tractography NOS is significantly (p\textless 0.05) correlated with ratings. However unfiltered tractography show no significance.}
\label{fig:SGfigure7}
\end{figure}

@@

When attempting to judge tractography delineation quality, human experts often are able to assess a particular delineation by visually identifying known anatomical priors. Direct tractography visualizations suggest that wide streamline dispersions often obscure anatomical details in visualizations, making anatomical assessment difficult (Figure \ref{fig:SGfigure3}). Therefore, image conjunctions offer better ability to assess reproducibility than direct tractography renderings alone. We aimed to develop an automated assessment of tractography reproducibility, since a) researchers are often prone to judgment bias when exposed to large number of tractography results, and that subtle differences in results are often hard to distinguish, and b) in an automated pipeline, the assessment score is necessary to allow iteratively fine tuning of tractography parameters. Since current neuroimaging techniques are not yet able to reliably classify neuroanatomy from tractography or image-based morphology, it is desirable for a computing pipeline to optimize and present the most reproducible tractography delineating across a population based on a particular set of ROIs as a first step. Subsequent anatomical assessment can then be made by a human expert, and the appropriate changes in ROI filter strategy can be made. This approach allows researchers with strong neuroanatomy backgrounds, but are less technically inclined, to make better judgments when performing group tractography.

@@

The limit of the NOS score is closed tied to the performance of its associated tractography algorithm. An algorithm and its associated parameters may consistently produce the wrong result in all subjects and produce a high NOS score, or it may produce no results, and result in a NOS score of 0. Both of these cases are regular occurrences in single-subject tractography analysis. With merged tractography, it is easier to recognize such faults. This is because in practice, incorrect tract path are often the result of imaging anomalies or algorithm limit, and are unstable across a population. This can be observed as a general trend in Figure \ref{fig:SGfigure3}, for example where DTI delineation of the rubrocerebellar tract is clearly incorrect, and results in low NOS score. Therefore, NOS score is capable of characterizing the uncertainty as a result of low algorithmic performance. In the case of false positives that are highly consistent, SAGIT can help researchers making more informed judgement.
